---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# animbook

<!-- badges: start -->
<!-- badges: end -->

"animbook" is a package to help the user visualize the changes in performance measures and demographic affiliations using animation. It is a package to help prepare, plot, and animate the data.


## Installation

You can install the development version of animbook from [GitHub](https://github.com/KrisanatA/animbook) with:

``` r
# install.packages("devtools")
devtools::install_github("KrisanatA/animbook")
```

## Usage

```{r}
library(animbook)
```

We will be using the dataset provided by the package `osiris`. There are three stages in this package that the users need to follow.


### First stage: Prepare the data

In this stage, we first need to prepare the data structure to a specific format. This function will assign a frame, calculate the scaling, and create the necessary data and settings for the next stage.

The data needs to contain the following variable for the function to work:

- id: `ID`, it needs to be a factor variable.
- values: `sales`, needs to be a numeric variable.
- time: `year`, it needs to be an integer variable.

```{r}
animbook <- anim_prep(osiris, ID, sales, year)
```

There are also additional options that allow the user to customize.

- label: group labeling.
- ngroup: number of groups we want to split the values into.
- breaks: the group bins size (prototype)
- group_scaling: the grouping variable for the bins calculations.
- color: the variable used to color the observations.
- time_dependent: logical. Whether we want the observations to start at the same time or not.
- scaling: the scaling method, either `rank` or `absolute`.
- runif_min: minimum value for random addition to frame numbers
- runif_max: maximum value for random addition to frame numbers

The function can calculate four different scales using these options.

```{r}
# rank scaling
rank_scaling <- anim_prep(data = osiris, id = ID, values = sales, time = year)

# absolute scaling
absolute_scaling <- anim_prep(data = osiris, id = ID, values = sales, time = year,
                              scaling = "absolute")

# rank scaling by group
rank_group_scaling <- anim_prep(data = osiris, id = ID, values = sales, time = year, 
                                group = country)

# absolute scaling by group
absolute_group_scaling <- anim_prep(data = osiris, id = ID, values = sales, time = year,
                                    group = country, scaling = "absolute")

rank_scaling
```

This function will return an `animbook` object containing a list of the formatted data and settings.


### Second stage: Plotting the data

After preparing the data, we can not plot it. There are three plots available in this package:

- `kangaroo`, which plots the observation's movement over time.
- `wallaby`, which subset the plot to either `top` or `bottom` and see which group they are in after the observational period.
- `funnel_web_spider`, which is a faceted plot by time variable.

```{r}
label <- c("Top 20%", "20-40", "40-60", "60-80", "80-100", "not listed")

animbook <- anim_prep(data = osiris, id = ID, values = sales, time = year, color = japan, label = label)

# kangaroo plot
kangaroo_plot(animbook)

# wallaby plot
wallaby_plot(animbook)
            
# funnel web spider plot
funnel_web_plot(animbook)
```

The `kangaroo` and `wallaby` plots can be animated using the function of the next stage. `funnel_web_spider` only supported static plot. We can also choose whether we want to animate the plot using gganimate or plotly.


### Third stage: Animating the plot

To animate the plot, we need to save the plot into an object, which then can be passed on to the function.

```{r}
animbook <- anim_prep_cat(data = aeles, id = id, values = party, time = year, color = gender, time_dependent = FALSE)

p <- wallaby_plot(animbook)

p2 <- anim_animate(p)

gganimate::animate(p2)
```








